---
title: "Read In Velocyto and 10x objects for Maya"
output: 
  github_document:
    toc: true
  html_notebook:
    toc: true
---

# Environment

```{r}
library(tidyverse)
library(Seurat)
library(SeuratWrappers)
library(SeuratDisk)
library(velocyto.R)
```
Names of Maya's samples
```{r}
names <- c('OE_control','OE_IL13','OE_M2','OE_M2_IL13','OE_M4')
```
# Read in Velocyto loom files
Read in all objects as Velocity, convert to seurat, return list of objects.
There is some kind of bug in ReadVelocity where it is pulling gene symbols, which are not guaranteed unique. see issue #78. Will overwrite read.loom.matrices with a function that appends accession numbers to gene symbols to ensure uniqueness.
```{r}
read.loom.matrices <- function(file) {
  # modified from velocytoR source code to account for non-uniqueness.
  cat('reading loom file via hdf5r...\n')
    f <- hdf5r::H5File$new(file, mode='r')
    cells <- f[["col_attrs/CellID"]][]
    genes <- paste0(f[["row_attrs/Gene"]][],'.',f[["row_attrs/Accession"]][]) # changed to gene.accession
    dl <- c(spliced="layers/spliced",
            unspliced="layers/unspliced",
            ambiguous="layers/ambiguous")
    if("layers/spanning" %in% hdf5r::list.datasets(f)) {
      dl <- c(dl, c(spanning="layers/spanning"))
    }
    dlist <- lapply(dl, function(path) {
      m <- as(t(f[[path]][,]),'dgCMatrix')
      rownames(m) <- genes; colnames(m) <- cells;
      return(m)
    })
    f$close_all()
    return(dlist)
}
```

Actually read in the looms and convert to list of seurat objects. Also add some basic group metadata
```{r}
seu.velocity.list <- lapply(names, function(name){
  dir <- paste0('./count/',name,'/velocyto/')
  # Assumes a single loom file per library
  loom.filename <- list.files(dir, pattern='*.loom')
  vel.object <- read.loom.matrices(file=paste0(dir,loom.filename))
  seu.object <- as.Seurat(x=vel.object)
  # Add metadata
  seu.object$group <- name
  return(seu.object)
})
```

# Read in 10X data
```{r}
seu.10x.list <- lapply(names, function(name){
  dir <- paste0('./count/',name,'/outs/')
  raw.matrix <- Read10X(dir)
  seu.object <- CreateSeuratObject(counts=raw.matrix)
  # add metadata
  seu.object$group <- name
  return(seu.object)
})
```

# Merge velocity data with corresponding 10x data

## helper functions

### Function to rename velocity/loom cells
Velocity/Loom seems to have named cells in the form library:Barcodex, e.g., KO:AAACCCATCCCTCTCCx. Seurat appends a dash for whatever reason i suppose in case of nonuniqueness which is actually relatively improbable.
```{r}
ConvertVelocityCellToSeuratCell <- function(vector) {
  list <- lapply(vector, function(name) {
    name <- strsplit(name, ':')[[1]][2]
    name <- substr(name,1,nchar(name)-1)
    name <- paste0(name,'-1')
    return(name)
  })
  return(unlist(list))
}
```

### Helper function to transfer assays
Three assays come in the loom file: spliced, unspliced, and ambiguous. 
```{r}
ImportVelocityAssays <- function(Seurat10xObj, SeuratVelocityObj) {
  # Requires that the cell names have been matched already
  # Returns an object with all the data from both
  Seurat10xObj[['spliced']] <- SeuratVelocityObj[['spliced']]
  Seurat10xObj[['unspliced']] <- SeuratVelocityObj[['unspliced']]
  Seurat10xObj[['ambiguous']] <- SeuratVelocityObj[['ambiguous']]
  return(Seurat10xObj)
}
```

## Validate barcode match
One predicts that the converted names for the loom-Seurat objects will be identical to names in the Seurat10x objects. If this is not the case there is a problem.
```{r}
for(i in 1:length(names)){
  seu.10x.obj <- seu.10x.list[[i]]
  seu.vel.obj <- seu.velocity.list[[i]]
  print(names[i])
  print(setequal(ConvertVelocityCellToSeuratCell(Cells(seu.vel.obj)), Cells(seu.10x.obj)))
  rm(seu.vel.obj)
  rm(seu.10x.obj)
}
```

## Rename velocity cells

```{r}
seu.vel.renamed.list <- lapply(seu.velocity.list, function(seu.vel.obj){
  seu.vel.renamed.obj <- RenameCells(seu.vel.obj,
                                     new.names=ConvertVelocityCellToSeuratCell(Cells(seu.vel.obj)))
})
```

## Transfer assays to 10x objects
```{r}
seu.10x.vel.list <- lapply(1:length(names), function(i){
  seu.10x.obj <- seu.10x.list[[i]]
  seu.vel.renamed.obj <- seu.vel.renamed.list[[i]]
  seu.10x.vel.obj <- ImportVelocityAssays(seu.10x.obj, seu.vel.renamed.obj)
  # Set default assay in case it's ambiguous
  DefaultAssay(seu.10x.vel.obj) <- 'RNA'
  return(seu.10x.vel.obj)
})
```

## cleanup cleanup
```{r}
rm(seu.10x.list, seu.vel.renamed.list, seu.velocity.list)
gc()
```

# Merge and Integrate
Using the newer Seurat5 workflow which differs from previous analysis. 
## Merge
```{r}
# Seurat merge expects 1 object and a list of remainder.
tuft.vel.cells <- merge(seu.10x.vel.list[[1]], seu.10x.vel.list[-1])
```

## preprocessing
```{r}
tuft.vel.cells[['percent.mt']] <-  PercentageFeatureSet(tuft.vel.cells, pattern = "^mt-")
```
```{r}
VlnPlot(tuft.vel.cells, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0)
```
```{r}
tuft.vel.cells <- subset(tuft.vel.cells, subset = nFeature_RNA > 1000 & nFeature_RNA < 7500 & percent.mt < 7.5)
VlnPlot(tuft.vel.cells, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0)
```
## processing
```{r}
# general processing
tuft.vel.cells <- NormalizeData(tuft.vel.cells, verbose=F)
tuft.vel.cells <- FindVariableFeatures(tuft.vel.cells, verbose=F)
tuft.vel.cells <- ScaleData(tuft.vel.cells, features=rownames(tuft.vel.cells), verbose=F)
tuft.vel.cells <- RunPCA(tuft.vel.cells, verbose=F)
```

## integration
This is a new workflow for me.
```{r}
tuft.vel.cells <- IntegrateLayers(object=tuft.vel.cells, method=FastMNNIntegration,
                                  orig.reduction = "pca", new.reduction = "integrated.mnn",
                                  verbose = FALSE)
# "rejoin layers" says Seurat
tuft.vel.cells[["RNA"]] <- JoinLayers(tuft.vel.cells[["RNA"]])
```

## Clustering and UMAP embedding on integrated data
```{r}
tuft.vel.cells <- FindNeighbors(tuft.vel.cells, reduction='integrated.mnn', dims=1:30, verbose=F)
tuft.vel.cells <- FindClusters(tuft.vel.cells, resolution=0.5, verbose=F)
tuft.vel.cells <- RunUMAP(tuft.vel.cells, reduction='integrated.mnn', dims=1:30, verbose=F)
```
```{r}
DimPlot(tuft.vel.cells)
DimPlot(tuft.vel.cells, group.by = 'group')
DimPlot(tuft.vel.cells, split.by = 'group')
```

# Quick look

## Cell cycle
Did not scale data / adjust for cell cycle during pre-processing, add those scores here anyway.
```{r}
s.genes <- str_to_title(cc.genes$s.genes)
g2m.genes <- str_to_title(cc.genes$g2m.genes)
tuft.vel.cells <- CellCycleScoring(tuft.vel.cells, s.features = s.genes, g2m.features = g2m.genes, set.ident = F)
```
```{r}
FeaturePlot(tuft.vel.cells, features=c('S.Score','G2M.Score','nFeature_RNA','percent.mt'))
```


## Cluster exploration
```{r}
FeaturePlot(tuft.vel.cells, features=c('Pou2f3','Trpm5','Krt5','Krt8'))
FeaturePlot(tuft.vel.cells, features=c('Ptprc','Ccr2','Omp','Pdgfra'))
```

## Save object
```{r, eval=F}
save(tuft.vel.cells, file='tuft.vel.cells.Robj')
```


# VelocytoR
see if this makes any sense at all

## Velocity on OE_M2
When some of the action might actually be happening?
```{r}
bm <- RunVelocity(object = subset(tuft.vel.cells, subset=group=='OE_M2'), deltaT = 1, kCells = 25, fit.quantile = 0.02)
ident.colors <- (scales::hue_pal())(n = length(x = levels(x = bm)))
names(x = ident.colors) <- levels(x = bm)
cell.colors <- ident.colors[Idents(object = bm)]
names(x = cell.colors) <- colnames(x = bm)
```
```{r, fig.height=8, fig.width=10}
show.velocity.on.embedding.cor(emb = Embeddings(object = bm, reduction = "umap"), vel = Tool(object = bm, 
    slot = "RunVelocity"), n = 200, scale = "sqrt", cell.colors = ac(x = cell.colors, alpha = 0.5), 
    cex = 0.8, arrow.scale = 3, show.grid.flow = TRUE, min.grid.cell.mass = 0.5, grid.n = 40, arrow.lwd = 1, 
    do.par = FALSE, cell.border.alpha = 0.1)
```

